### 2024-10-30 kozlovskii_dobush_romanik_ujp_UTF8.tex
brushed for UJP, add ukrainian parts, add bibliography as bibitems, figures width=column

\begin{document}
	\title[Supercritical crossover lines in a cell fluid model]%колонтитул
	{Supercritical crossover lines in a cell fluid model}%
	\author{M.P.~Kozlovskii}%1
	\author{O.A.~Dobush}
	\author{R.V.~Romanik}
	%\author{I.V.~Pylyuk}
	\affiliation{Institute for Condensed Matter Physics,\\ Nat. Acad. of Sci. of Ukraine}%інститут 1
	\address{1, Svientsitskii Str., Lviv 79011, Ukraine}%адреса 1
	\email{mpk@icmp.lviv.ua}%e-mail 1
	\email{dobush@icmp.lviv.ua}
	\email{romanik@icmp.lviv.ua}
	%\email{piv@icmp.lviv.ua}

@@

	\begin{abstract}
		A cell fluid model with a modified Morse potential is studied. The supercritical states are considered with respect to a possibility to build a separation boundary between liquid-like and gas-like bahaviors. Three different lines are calculated that can be used for this purpose: the locus of the isothermal compressibility maxima, the locus of the thermal expansion coefficient maxima, and the line where the effective chemical potential is zero, $M=0$. By the symmetry of the functionals for the partition functions, the condition $M=0$ in fluids is analogous to the absence of an external field in the Ising model.
	\end{abstract}

	\keywords{Cell model, Morse potential, Thermodynamic response functions}

@@

	\input{./paper_text.tex}
	\bibliographystyle{gost2008}
	\bibliography{extracted}
	%\bibliography{extracted}

	\vspace*{-5mm}\rezume{%Резюме укр. мовою
		М.П.~Козловський, О.А.~Добуш, Р.В.~Романік}{Надкритичні лінії переходу в комірковій моделі плину} {В роботі проведено... .}{\textit{К\,л\,ю\,ч\,о\,в\,і\,
			с\,л\,о\,в\,а:} коміркова модель, потенціал Морзе, термодинамічні функції відгуку.}

>>>>

\begin{document}
	\title[Supercritical crossover lines in a cell fluid model]%колонтитул
	{Supercritical crossover lines in a cell fluid model}%
	\author{O.A.~Dobush}
	\author{M.P.~Kozlovskii}%1
	\author{R.V.~Romanik}
	%\author{I.V.~Pylyuk}
	\affiliation{Institute for Condensed Matter Physics,\\ Nat. Acad. of Sci. of Ukraine}%інститут 1
	\address{1, Svientsitskii Str., Lviv 79011, Ukraine}%адреса 1
	\email{mpk@icmp.lviv.ua}%e-mail 1
	\email{dobush@icmp.lviv.ua}
	\email{romanik@icmp.lviv.ua}
	%\email{piv@icmp.lviv.ua}

@@

	\begin{abstract}
		A cell fluid model with a modified Morse potential is studied. The supercritical states are considered with respect to the possibility of building a separation boundary between liquid-like and gas-like behaviors. Three different lines are calculated that can be used for this purpose: the locus of the isothermal compressibility maxima, the locus of the thermal expansion coefficient maxima, and the line where the effective chemical potential is zero, $M=0$. By the symmetry of the functionals for the partition functions, the condition $M=0$ in fluids is analogous to the absence of an external field in the Ising model.
	\end{abstract}

	\keywords{Cell fluid model, Morse potential, Supercritical region, Widom line}

@@

	\input{./paper_text.tex}
	%\bibliographystyle{gost2008}
	%\bibliography{extracted}
	%\bibliography{extracted}

	\vspace*{-5mm}\rezume{%Резюме укр. мовою
		М.П.~Козловський, О.А.~Добуш, Р.В.~Романік}{Надкритичні лінії переходу в комірковій моделі плину}{Продовжується дослідження коміркової моделі плину з модифікованим потенціалом Морзе. Надкритичні стани розглядаються з огляду на можливість побудови межі розділення між рідиноподібними та газоподібними поведінками. Розраховано три різні лінії, які можна використовувати для цієї мети: лінія максимумів ізотермічної стисливості, лінія максимумів коефіцієнта теплового розширення та лінія, де ефективний хімічний потенціал дорівнює нулю, $M=0$. За симетрією функціоналів статистичних сум умова $M=0$ у рідинах аналогічна відсутності зовнішнього поля в моделі Ізінга.}{\textit{К\,л\,ю\,ч\,о\,в\,і\, с\,л\,о\,в\,а:} коміркова модель, потенціал Морзе, надкритична область, лінія Відома.}

### 2024-10-30 paper_text.tex
brushed for UJP, add ukrainian parts, add bibliography as bibitems, figures width=column

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f1.pdf} 
	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$. Black lines correspond to $T \geq T_c$. Grey lines correspond to $T < T_c$. 
	}\label{fig1}
\end{figure}
Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various relative temperature values $\tau$. Below the critical temperature ($\tau < 0$), each іsotherm contains a kink (three bottom lines in Figure~\ref{fig1}). This form indicates a first-order phase transition in the system. Within our approach~\cite{KD20} at $T<T_c$, negative values of the effective chemical potential $M$ imply the system is in a gaseous state. The condition $M=0$ is a point of phase coexistence. Positive values of the effective chemical potential correspond to the liquid state. In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much steeper. At the critical temperature, the pressure is a smooth curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$. Thus, at the critical point coordinates are defined by $M=0, \tau=0$, $P^*_c = 1.606$. Beyond the critical temperature, the pressure behaves as a monotonically increasing function, and no phase transitions are observed. Summing up, below the critical temperature, $M=0$ is the condition for phase coexistence. Substituting this condition into the EoS~\eqref{eq:eosMT}, we obtain a phase diagram in pressure-temperature plane (see Figure~\ref{fig2}). We will use the same condition to obtain a continuation of the coexistence line into the supercritical region. 
\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f2.pdf} 
	\vskip-3mm\caption{Phase diagram in the $P^*-\tau$ plane for the cell fluid model with the interaction potential \eqref{def:mod_morse}. The result is based on the equation of state~\eqref{eq:eosMT}. 
	}\label{fig2}
\end{figure}

@@

	%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
	\centering \includegraphics[width=0.46\textwidth]{f3bb.pdf}
	%\includegraphics[width=\column]{f2b.pdf}
	%\includegraphics[width=\column]{f2c.pdf}

@@

	%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
	\centering \includegraphics[width=0.46\textwidth]{f3cc.pdf}
	%\includegraphics[width=\column]{f2b.pdf}
	%\includegraphics[width=\column]{f2c.pdf}

@@

\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f4b.pdf}
	\vskip-3mm\caption{Complete pressure-temperature phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}. Comparison of the Widom line (maxima locus of isothermal compressibility and thermal pressure) and the crossover between gas-like and liquid-like structures of the supercritical cell fluid based on combination of the equation of state~\eqref{eq:eosMT} and the condition $M = 0$. An inset in the right bottom corner focuses on a range of $\tau$ around its critical value $-0.01 < \tau < 0.05$.	}\label{fig4}
\end{figure}

@@

>>>>

\begin{figure}[t!]
	\centering
	\includegraphics[width=\column]{f1.pdf} 
	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$. Black lines correspond to $T \geq T_c$. Grey lines correspond to $T < T_c$. 
	}\label{fig1}
\end{figure}
Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various relative temperature values $\tau$. Below the critical temperature ($\tau < 0$), each іsotherm contains a kink (three bottom lines in Figure~\ref{fig1}). This form indicates a first-order phase transition in the system. Within our approach~\cite{KD20} at $T<T_c$, negative values of the effective chemical potential $M$ imply the system is in a gaseous state. The condition $M=0$ is a point of phase coexistence. Positive values of the effective chemical potential correspond to the liquid state. In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much steeper. At the critical temperature, the pressure is a smooth curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$. Thus, at the critical point coordinates are defined by $M=0, \tau=0$, $P^*_c = 1.606$. Beyond the critical temperature, the pressure behaves as a monotonically increasing function, and no phase transitions are observed. Summing up, below the critical temperature, $M=0$ is the condition for phase coexistence. Substituting this condition into the EoS~\eqref{eq:eosMT}, we obtain a phase diagram in pressure-temperature plane (see Figure~\ref{fig2}). We will use the same condition to obtain a continuation of the coexistence line into the supercritical region. 
\begin{figure}[t!]
	\centering
	\includegraphics[width=\column]{f2.pdf} 
	\vskip-3mm\caption{Phase diagram in the $P^*-\tau$ plane for the cell fluid model with the interaction potential \eqref{def:mod_morse}. The result is based on the equation of state~\eqref{eq:eosMT}. 
	}\label{fig2}
\end{figure}

@@

	\centering \includegraphics[width=\column]{f3bb.pdf}

@@

	\centering \includegraphics[width=\column]{f3cc.pdf}

@@

\begin{figure}[h!]
	\centering \includegraphics[width=\column]{f4b.pdf}
	\vskip-3mm\caption{Complete pressure-temperature phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}. Comparison of the Widom line (maxima locus of isothermal compressibility and thermal pressure) and the crossover between gas-like and liquid-like structures of the supercritical cell fluid based on combination of the equation of state~\eqref{eq:eosMT} and the condition $M = 0$. An inset in the right bottom corner focuses on a range of $\tau$ around its critical value $-0.01 < \tau < 0.05$.	}\label{fig4}
\end{figure}

@@

\begin{thebibliography}{99}

	%1
	\bibitem{KMLetal14} {\v{Z}}.~Knez, E.~Marko{\v{c}}i{\v{c}}, M.~Leitgeb, M.~Primo{\v{z}}i{\v{c}}, M.~Knez Hrn{\v{c}}i{\v{c}}, M.~{\v{S}}kerget.
	\href{https://doi.org/10.1016/j.energy.2014.07.044}{Industrial applications of supercritical fluids: A review}. \textit{Energy} \textbf{77} 235 (2014) [doi: 10.1016/j.energy.2014.07.044]
	%2
	\bibitem{FW69} M.E.~Fisher, B.~Widom.
	\href{https://doi.org/10.1063/1.1671624}{Decay of Correlations in Linear Systems}. \textit{J. Chem. Phys.} \textbf{50} 3756 (1969) [doi: 10.1063/1.1671624]
	%3
	\bibitem{BGMetal17} T.~Bryk, F.A.~Gorelli, I.~Mryglod, G.~Ruocco, M.~Santoro, T.~Scopigno.
	\href{https://doi.org/10.1021/acs.jpclett.7b02176}{Behavior of Supercritical Fluids across the “Frenkel Line”}. \textit{J. Chem. Phys. Lett.} \textbf{8} 4995 (2017) [doi: 10.1021/acs.jpclett.7b02176]
	%4
	\bibitem{CKS01} X.~Campi, H.~Krivine, N.~Sator.
	\href{https://doi.org/10.1016/S0378-4371(01)00158-3}{Percolation line of self-bound clusters in supercritical fluids}. \textit{Physica A} \textbf{296} 24 (2001) [doi: 10.1016/S0378-4371(01)00158-3] 
	%5
	\bibitem{XKBetal05} L.~Xu, P.~Kumar, S.V.~Buldyrev, S.-H.~Chen, P.H.~Poole, F.~Sciortino, H.E.~Stanley.
	\href{https://doi.org/10.1073/pnas.0507870102}{Relation between the Widom line and the dynamic crossover in systems with a liquid-liquid phase transition}. \textit{PNAS} \textbf{102} 16558 (2005) [doi: 10.1073/pnas.0507870102]
	%6
	\bibitem{MS10} P.~McMillan, H.~Stanley.
	\href{https://doi.org/10.1038/nphys1711}{Going supercritical}. \textit{Nature Phys.} \textbf{6} 479 (2010) [doi: 10.1038/nphys1711]
	%7
	\bibitem{AKM17} S.~Artemenko, P.~Krijgsman, V.~Mazur. 
	\href{https://doi.org/10.1016/j.molliq.2017.03.107}{The Widom line for supercritical fluids}. \textit{J. Mol. Liq.} \textbf{238} 122 (2017) [doi: 10.1016/j.molliq.2017.03.107]
	%8
	\bibitem{Proctor20} J.E.~Proctor.
	\href{https://doi.org/10.1201/9780429491443}{The liquid and supercritical fluid states of matter}. (CRC Press, 2020)  [ISBN: 9780429491443]
	%9
	\bibitem{KD20} M.~Kozlovskii, O.~Dobush.
	\href{https://ujp.bitp.kiev.ua/index.php/ujp/article/view/2019631}{Phase behavior of a cell fluid model with modified Morse potential.} \textit{Ukr. J. Phys.} \textbf{65}, 428 (2020) [doi: 10.15407/ujpe65.5.428]
	%10
	\bibitem{DKRP24} O.A.~Dobush, M.P.~Kozlovskii, R.V.~Romanik, I.V.~Pylyuk.
	\href{https://doi.org/10.48550/arXiv.2409.09786}{Thermodynamic response functions in a cell fluid model}. \textit{Ukr. J. Phys} - in print [doi: 10.48550/arXiv.2409.09786]
	%11
	\bibitem{DKRP24arxiv} O.A.~Dobush, M.P.~Kozlovskii, R.V.~Romanik, I.V.~Pylyuk.
	\href{https://doi.org/10.48550/arXiv.2409.09786}{Thermodynamic response functions in a cell fluid model}. \textit{arXiv preprint} 2409.09786 (2024) [doi: 10.48550/arXiv.2409.09786]
	%12
	\bibitem{Morse29} P.~M.~Morse. 
	\href{https://link.aps.org/doi/10.1103/PhysRev.34.57}{Diatomic molecules according to the wave mechanics. II. Vibrational levels.} \textit{Phys. Rev.} \textbf{34}, 57 (1929) [doi: 10.1103/PhysRev.34.57]
	%13
	\bibitem{SBGetal10} G.G.~Simeoni, T.~Bryk, F.A.~Gorelli, M.~Krisch, G.~Ruocco, M.~Santoro, T.~Scopigno.
	\href{https://doi.org/10.1038/nphys1683}{ The Widom line as the crossover between liquid-like and gas-like behaviour in supercritical fluids}. \textit{Nature Phys.} \textbf{6} 503 (2010) [doi: 10.1038/nphys1683]
	%14
	\bibitem{KozlovskiiDobush2016} M.~Kozlovskii, O.~Dobush.
	\href{https://doi.org/j.molliq.2015.12.018}{Representation of the grand partition function of the cell model: The state equation in the mean-field approximation.} \textit{J. Mol. Liq.} \textbf{215}, 58 (2016) [doi: j.molliq.2015.12.018]
	%15
	\bibitem{KR09} M.P~Kozlovskii, R.V.~Romanik.
	\href{https://doi.org/10.30970/jps.13.4007}{The equation of state of a three-dimensional {Ising}-like system}. \textit{Journal of Physical Studies} \textbf{13} 4007 (2009) [doi: 10.30970/jps.13.4007]
	%16
	\bibitem{Kulinskii10jcp} V.L.~Kulinskii.
	\href{https://doi.org/10.1063/1.3457943}{Global isomorphism between the {Lennard-Jones} fluids and the {Ising} model}. \textit{J. Chem. Phys.} \textbf{133} 034121 (2010) [doi: 10.1063/1.3457943]
	%17
	\bibitem{BK11} L.A.~Bulavin, V.L.~Kulinskii.
	\href{https://doi.org/10.1021/jp201872f}{Unified Picture for the Classical Laws of {B}atschinski and the Rectilinear Diameter for Molecular Fluids}. \textit{J. Phys. Chem. B} \textbf{115} 6061 (2011) [doi: 10.1021/jp201872f]

\end{thebibliography}
